SRO Retest Data Cleaning

### 2018-09-13 code/SRO_Retest_Cleaning.Rmd
reset repo

---
title: 'Self Regulation Ontology Retest Data Report'
output:
github_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(psych)
library(rmarkdown)
library(psych)
library(stringr)
library(plotly)
library(jsonlite)
library(DT)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('SRO_Report.Rmd', output_dir = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Report', html_notebook(toc = T, toc_float = T, code_folding = 'hide'))}
options(scipen = 1, digits = 4)
```

# Introduction

# Methods

## Data collection

This sample consists of data for 150 subjects of the original sample of 522 that has completed the initial battery of 37 cognitive tasks, 23 surveys and 3 surveys on demographics. Details of the original sample as well as quality control (qc) procedures are described elsewhere. Invited participants were chosen randomly and only subsets of them were invited for a given batch (instead of opening the battery to all qualified subjects) with the intention to avoid a potential oversampling and bias towards "high self regulators".

```{r warning=FALSE, message=FALSE}
workers = read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/Local/User_717570_workers.csv')
workers = workers %>% 
  group_by(Worker.ID) %>%
  mutate(Retest_worker=ifelse(sum(CURRENT.RetestWorker,CURRENT.RetestWorkerB2,CURRENT.RetestWorkerB3,CURRENT.RetestWorkerB4,CURRENT.RetestWorkerB5,na.rm=T)>0,1,0)) %>%
  ungroup()

worker_counts <- fromJSON('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/Local/retest_worker_counts.json')

worker_counts = as.data.frame(unlist(worker_counts))
names(worker_counts) = "task_count"
```

In total `r sum(workers$Retest_worker)` participants were invited, `r nrow(worker_counts)` began the battery, `r sum(worker_counts$task_count >= 62)` completed the battery and 150 provided data that passed qc for both time points. Our target sample size was determined in advance of data collection and data collection continued until this number of participants with data that survived qc was reached.

```{r warning=FALSE, message=FALSE}
disc_comp_date = read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/Local/discovery_completion_dates.csv', header=FALSE)
val_comp_date = read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/Local/validation_completion_dates.csv', header=FALSE)
test_comp_date = rbind(disc_comp_date, val_comp_date)
rm(disc_comp_date, val_comp_date)
retest_comp_date = read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/Local/retest_completion_dates.csv', header=FALSE)
comp_dates = merge(retest_comp_date, test_comp_date, by="V1")
names(comp_dates) <- c("sub_id", "retest_comp", "test_comp")
comp_dates$retest_comp = as.Date(comp_dates$retest_comp)
comp_dates$test_comp = as.Date(comp_dates$test_comp)
comp_dates$days_btw = with(comp_dates, retest_comp-test_comp)
```

Data collection took place on average `r round(mean(as.numeric(comp_dates$days_btw)))` number of days after the completion of the initial battery with a range of `r round(range(as.numeric(comp_dates$days_btw)))[1]` to `r round(range(as.numeric(comp_dates$days_btw)))[2]` days.

## Literature

### Tasks

### Surveys

## Loading and matching datasets

This report uses the variables that were designated as "meaningful" before. This is not the smallest subset where certain variables are dropped because they are correlated with other variables (within a task) and includes both accuracies as well as both kinds of DDM variables (EZ and HDDM). It will not include variables that were considered not of theoretical interest in the analyses of time 1 data. The variables that are not included will be listed for reference as well.

### Load time 1 data
```{r}
test_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables.csv')

test_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_noDDM.csv')

test_data3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_EZ.csv')

test_data <- merge(test_data, test_data2)
test_data <- merge(test_data, test_data3)
rm(test_data2, test_data3)
```

For reference here are the variables that are **not** included in the analyses of the remainder of this report because they were not of theoretical interest in factor structure analyses of this data so far. These include drift diffusion and other model parameters for specific conditions within a task; survey variables that are not part of the dependant variables for that survey in the literature and demographics (these are saved for prediction analyses).

```{r}
test_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/variables_exhaustive.csv')

test_data2$X <- as.character(test_data2$X)

names(test_data2)[1] <- 'sub_id'

df <- data.frame(names(test_data2)[which(names(test_data2) %in% names(test_data) == FALSE)])
names(df) = c('vars')

datatable(df)
```

```{r echo=FALSE}
rm(test_data2, df)
```

### Load time 2 data 
```{r}
retest_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/meaningful_variables.csv')

retest_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/meaningful_variables_noDDM.csv')

retest_data3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/meaningful_variables_EZ.csv')

retest_data <- merge(retest_data, retest_data2)
retest_data <- merge(retest_data, retest_data3)
rm(retest_data2, retest_data3)
```

Extract t2 participants with good t1 data: We only invited those with good data but due to a technical error a handful of subjects who weren't checked for data quality completed the retest battery as well. One of those subjects gave good t2 data but didn't have good t1 data and therefore is removed from further analyses.

```{r}
#Process sub_id columns in the data dataframes
retest_data$X <- as.character(retest_data$X)
test_data$X <- as.character(test_data$X)

names(retest_data)[1] <- 'sub_id'
names(test_data)[1] <- 'sub_id'

retest_data <- retest_data[retest_data$sub_id %in% test_data$sub_id,]
```

@@

### Extract t-1 data for t-2 subjects
```{r}
retest_subs_test_data <- test_data[test_data$sub_id %in% retest_data$sub_id,]

rest_subs_test_data <- test_data[test_data$sub_id %in% retest_data$sub_id == FALSE,]

#Arrange datasets of same size by sub_id
retest_data = retest_data %>% arrange(sub_id)
retest_subs_test_data = retest_subs_test_data %>% arrange(sub_id)

#CHECK IF EVERYTHING IS ORDERED RIGHT
as.character(retest_subs_test_data$sub_id) == as.character(retest_data$sub_id)
```

### Replace HDDM parameters in t1 data

Since HDDM parameters depend on the sample on which they are fit we refit the model on t1 data for the subjects that have t2 data. Here we replace the HDDM parameters in the current t1 dataset with these refitted values and transform them as they would have been transformed for the full sample to mimick the procedure we applied to rest of the t2 data.

Transform hddm refits as they had been transformed for the full sample t1 data.
```{r}
hddm_refits <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_03-21-2017/hddm_refits_exhaustive.csv')

hddm_refits$X <- as.character(hddm_refits$X)

names(hddm_refits)[1] <- 'sub_id'

hddm_logTr_vars = data.frame(grep('(?=.*hddm)(?=.*logTr)', names(retest_subs_test_data), value=T, perl=T))
names(hddm_logTr_vars) = c('var')
hddm_logTr_vars$sign = ifelse(grepl(".ReflogTr",hddm_logTr_vars$var), 'negative', 'positive')
hddm_logTr_vars$var = gsub(".logTr|.ReflogTr", "", hddm_logTr_vars$var)

transformed_hddm_refits = hddm_refits

for(i in 1:length(names(transformed_hddm_refits))){
  tmp_col = names(transformed_hddm_refits)[i]
  if(tmp_col %in% hddm_logTr_vars$var){
    sign = hddm_logTr_vars$sign[which(tmp_col == hddm_logTr_vars$var)]
      if(sign == 'positive'){
        # clean_retest_data[,tmp_col] = log(clean_retest_data[,tmp_col])
        transformed_hddm_refits[,tmp_col] = pos_log(transformed_hddm_refits[,tmp_col])
        names(transformed_hddm_refits)[which(names(transformed_hddm_refits) == tmp_col)] = paste0(tmp_col, '.logTr')
        cat(paste0(tmp_col ,' was positively skewed and transformed.'))
        cat('\n')
      }
      else if(sign == 'negative'){
        transformed_hddm_refits[,tmp_col] = neg_log(transformed_hddm_refits[,tmp_col])
        names(transformed_hddm_refits)[which(names(transformed_hddm_refits) == tmp_col)] = paste0(tmp_col, '.ReflogTr')
        cat(paste0(tmp_col ,' was negatively skewed and transformed.'))
        cat('\n')
      }
    }
}
```

Check if transformation of hddm refits worked
```{r}
transform_hddm_logTr_vars = data.frame(grep('logTr', names(transformed_hddm_refits), value=T))
names(transform_hddm_logTr_vars) = c('var')
transform_hddm_logTr_vars$sign = ifelse(grepl(".ReflogTr",transform_hddm_logTr_vars$var), 'negative', 'positive')
transform_hddm_logTr_vars$var = gsub(".logTr|.ReflogTr", "", transform_hddm_logTr_vars$var)
transform_hddm_logTr_vars = transform_hddm_logTr_vars %>% arrange(var)
hddm_logTr_vars = hddm_logTr_vars %>% arrange(var)
transform_hddm_logTr_vars == hddm_logTr_vars
```

Replace t1 hddm parameters for retest subjects

```{r}
replace_hddm_vars = data.frame(grep('(?=.*hddm)(?=.*logTr)', names(retest_subs_test_data), value=T, perl=T))
names(replace_hddm_vars) = c('var')

hddm_fix_retest_subs_test_data = retest_subs_test_data

for(i in 1:length(names(hddm_fix_retest_subs_test_data))){
  if(names(hddm_fix_retest_subs_test_data)[i] %in% replace_hddm_vars$var){
    hddm_fix_retest_subs_test_data[,names(hddm_fix_retest_subs_test_data)[i]] <- transformed_hddm_refits[,names(hddm_fix_retest_subs_test_data)[i]]
  }
}

retest_subs_test_data <- hddm_fix_retest_subs_test_data
```

```{r}
rm(hddm_refits, transformed_hddm_refits, hddm_logTr_vars, transform_hddm_logTr_vars, replace_hddm_vars, hddm_fix_retest_subs_test_data, sign, i)
```

# Results

## Reliability metrics 
For correlations we report Spearman's $\rho$'s only.  
Based on [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) ICC(3,k) does not take in to account within subject differences between two time points (i.e. the fixed effect of time/systematic error). Thus, it is well approximated by Pearson's r and subject to similar criticisms. [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) suggests reporting at least this systematic error effect size if one chooses to report with ICC(3,k). Based on his conclusions here I report:  
- ICC(3,k): As Dave clarified this ranges from 1 to -1/(number of repeated measures -1) so in our case this range would be [-1, 1]; larger values would mean that the two scores of a subject for a given measure are more similar to each other than they are to scores of other people  
- partial $\eta^2$ for time ($SS_{time}/SS_{within}$): effect size of time   
- SEM ($\sqrt(MS_{error})$): standard error of measurement; the smaller the better

```{r}
numeric_cols = c()

match_t1_t2 <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', format = "long", sample = 'full', sample_vec){

  if(sample == 'full'){
   df = merge(t1_df[,c(merge_var, dv_var)], t2_df[,c(merge_var, dv_var)], by = merge_var) 
  }
  else{
    df = merge(t1_df[t1_df[,merge_var] %in% sample_vec, c(merge_var, dv_var)], t2_df[t2_df[,merge_var] %in% sample_vec, c(merge_var, dv_var)],
    by=merge_var)
  }

  df = df %>% 
    na.omit()%>%
    gather(dv, score, -sub_id) %>%
    mutate(time = ifelse(grepl('\\.x', dv), 1, ifelse(grepl('\\.y', dv), 2, NA))) %>%
    separate(dv, c("dv", "drop"), sep='\\.([^.]*)$') %>%
    select(-drop)

  if(format == 'wide'){
    df = df%>% spread(time, score) 
  }

  return(df)
}

get_spearman = function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){

  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide')
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide', sample='bootstrap', sample_vec = sample_vec)
  }

  rho = cor(df$`1`, df$`2`, method='spearman')

  return(rho)
}

get_icc <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide')
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide', sample='bootstrap', sample_vec = sample_vec)
  }

  df = df %>% select(-dv, -sub_id)
  icc = ICC(df)
  icc_3k = icc$results['Average_fixed_raters', 'ICC']
  return(icc_3k)
  }

get_eta <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, sample='bootstrap', sample_vec = sample_vec)
  }

  mod = summary(aov(score~Error(sub_id)+time, df))
  ss_time = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq1',]
  ss_error = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq2',]
  eta = ss_time/(ss_time+ss_error)
  return(eta)
  }

get_sem <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, sample='bootstrap', sample_vec = sample_vec)
  }
  mod = summary(aov(score~Error(sub_id)+time, df))
  ms_error = as.data.frame(unlist(mod$`Error: Within`))['Mean Sq2',]
  sem = sqrt(ms_error)
  return(sem)
}
```

```{r}
rel_df <- data.frame(spearman = rep(NA, length(numeric_cols)),
                     icc = rep(NA, length(numeric_cols)), 
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)))

row.names(rel_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  rel_df[numeric_cols[i], 'spearman'] <- get_spearman(numeric_cols[i]) 
  rel_df[numeric_cols[i], 'icc'] <- get_icc(numeric_cols[i])
  rel_df[numeric_cols[i], 'eta_sq'] <- get_eta(numeric_cols[i])
  rel_df[numeric_cols[i], 'sem'] <- get_sem(numeric_cols[i])
}

rel_df$dv = row.names(rel_df)
row.names(rel_df) = seq(1:nrow(rel_df))
rel_df$task = 'task'
rel_df[grep('survey', rel_df$dv), 'task'] = 'survey'
rel_df = rel_df %>%
  select(dv, task, spearman, icc, eta_sq, sem)
# row.names(rel_df) = NULL

```

Average reliability metrics and the number of variables for both the surveys and the tasks
Both Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r warning=FALSE, message=FALSE}
rel_df %>%
  group_by(task) %>%
  summarise(median_spearman = median(spearman),
            median_icc = median(icc),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n()) %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc', 'median_eta_sq', 'median_sem'), digits=3)
```

*Rank orders for reliabilities*

- Note that this table can be sorted by different metrics, filtered for specific values for each column and is searchable.  
- Demographics are excluded from this list of variables (because they are saved for predictive analyses).  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to our tasks (one that is used more often by economists). In previous data I have worked with retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.

```{r}
rel_df %>%
  datatable(filter='top') %>%
  formatRound(columns=c('spearman', 'icc', 'eta_sq', 'sem'), digits=3)
```

### Histograms of reliability metrics

```{r echo=FALSE}
#Convenience functions to make prettier interactive histograms
get_labels = function(plot_df, ref_df, bin_col, group_col){

  plot_df$bin = as.numeric(factor(plot_df$x))
  ref_df$bin = NA

  for(i in 1:nrow(ref_df)){
    for(j in 1:nrow(plot_df)){
      # if(ref_df[i, bin_col] <= plot_df$xmax[j] & ref_df[i, bin_col]>=plot_df$xmin[j] & all(ref_df[i, group_col] == plot_df[j, group_col])){
      if(ref_df[i, bin_col] <= plot_df$xmax[j] & ref_df[i, bin_col]>=plot_df$xmin[j] & ref_df[i, group_col] == plot_df[j, group_col]){
        ref_df$bin[i] <- plot_df$bin[j]
      }
    }
  }

  ref_df = ref_df %>% 
    group_by_("bin", group_col) %>%
    mutate(label=as.character(toJSON(unique(dv))))

  label_df = ref_df[!duplicated(ref_df[,c('label')]),c(group_col, 'bin', 'label')]

  out_df = left_join(plot_df,label_df, by= c("bin", group_col))

  return(out_df)
}

line_break <- function(cell){
  out = paste(unlist(strsplit(gsub("\\]","",gsub("\\[","",cell)), ",")), sep="\n", collapse = "</br>")
  return(out)
}

prettify_label = function(df){
  for(i in 1:nrow(df)){
    df$label[i] =line_break(df$label[i])
  }
  return(df)
}
```

### Spearman

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(spearman, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "spearman", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Spearman')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

### ICC

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(icc, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "icc", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('ICC')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

### Partial $\eta^2$

Tooltip might not work well for this graph for the highest bar because the label is too large. But this bar indicates measures where the effect of time is very small anyway. We might want to pay closer attention to those not in this lowest bar since those are the variables where there is a non-negligible effect of time, indicating either a learning effect or other kind of systematic change that we did not hypothesize to observe with any of these measures.

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(eta_sq, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "eta_sq", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Partial Eta Squared')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

#### Effect of time

So what variables show a non-negligible effect of time and in what direction? (sorted by increasing partial $\eta^2$)

```{r warning=FALSE, message=FALSE}
time_effect_vars = ggplot_build(rel_df %>%
               ggplot(aes(eta_sq, fill=task)) +
               geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
  mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
  get_labels(.,rel_df, "eta_sq", "task") %>%
  prettify_label(.) %>%
  filter(bin != 1 & bin !=2 & label != "NA") %>%
  select(label, task) %>%
  separate(label,into=c(as.character(1:27)),sep = "</br>") %>%
  gather(key, value, -task) %>%
  select(value, task) %>%
  na.omit()

time_effect_vars$value = gsub("\"", "", time_effect_vars$value)
datatable(time_effect_vars)
```

Box plots showing the distributions of difference scores between the measurements at two time points for each measure can be found under the [Completion times](#diff_plot) section.

### SEM

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(sem, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "sem", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('SEM')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

### Negative icc's?
```{r}
neg_icc_vars = rel_df %>%
  arrange(icc) %>% 
  filter(icc<0) %>%
  select(dv)
```

Some variables (`r neg_icc_vars$dv`) have <0 ICC's. This would be the case if the $MS_{error}$>$MS_{between}$.

Plot data for measures with negative ICC.  
x-axis is time 1 data and y-axis is time 2 data. The scatter plots show that the scores for measures with negative ICC's have no relationship across time points.
```{r, fig.height=12, fig.width=12}
tmp = data.frame()

for(i in 1:nrow(neg_icc_vars)){
  tmp = rbind(tmp, match_t1_t2(neg_icc_vars$dv[i]))
}
rm(neg_icc_vars)

tmp %>%
  spread(time, score) %>%
  # mutate(dv_wrap = str_wrap(dv, width = 30)) %>%
  ggplot(aes(`1`, `2`))+
  geom_point()+
  theme_bw()+
  facet_wrap(~dv, scales='free', labeller = label_wrap_gen())

```

This point is clarified further by compare these scatter plots to those of variables with highest ICC. These variables have strong and clear positive linear relationships.

```{r fig.height=12, fig.width=12}
high_icc_vars = rel_df %>%
  arrange(-icc)  %>% 
  filter(icc>0.92) %>%
  select(dv)

tmp = data.frame()

for(i in 1:nrow(high_icc_vars)){
  tmp = rbind(tmp, match_t1_t2(high_icc_vars$dv[i]))
}
rm(high_icc_vars)

tmp %>%
  spread(time, score) %>%
  ggplot(aes(`1`, `2`))+
  geom_point()+
  theme_bw()+
  facet_wrap(~dv, scales='free')

```

```{r echo=FALSE}
rm(tmp)
```

### Relationship between ICC and Spearman
These two measures give mostly similar answers regarding the reliabilities of the measures except for

```{r}
ggplotly(rel_df %>%
           ggplot(aes(spearman, icc, col=task, label=dv, label2=eta_sq))+
           geom_point()+
           theme_bw()+
           theme(legend.title = element_blank())+
           geom_abline(intercept = 0, slope=1))
```

What are the variables clustered around 0 ICC?

```{r}
rel_df %>%
  filter(icc>(-0.1), icc<0.1) %>%
  arrange(icc) %>%
  datatable() %>%
  formatRound(columns=c('spearman', 'icc', 'eta_sq', 'sem'), digits=3)
```

### Relationship between ICC and $\eta^2$

```{r}
ggplotly(rel_df %>%
  ggplot(aes(eta_sq, icc, col=task, label=dv))+
  geom_point()+
  theme_bw()+
  theme(legend.title = element_blank()))
```

## Reliability by measure types

The results presented so far aim to be comprehensive but are difficult to parse through and get a clear sense of what is going on (other than a general impression that surveys tend to have higher reliabilities than tasks). In this and the following section we try to deal with this in two ways to make the results more digestable with two kinds of categorizations.

### Tasks

We aim to make contact with the literature to make the results more accesible to readers who have experience with our tasks but do not want to   
1. take away from the novelty of the main papers coming out of this project outlining the factor structure that this data unveils   
2. necessarily make prescriptive recommendations or theoretical arguments on why certain measures of certain processes are more reliable than others.

With this in mind we reduce the list of over 270 task measures to a list of one per task by averaging only the raw measures from all the trials in a task. We chose to reduce the information in this manner to avoid any bias stemming from differential amount of interest and procedures applied to certain tasks over others (e.g. a task can have over 10 measures because it has multiple conditions and we have chosen to fit DDM's for specific conditions while another might only have 2 due to our relative inexperience and lack of interest in it). We check whether the number of trials in a task has a significant effect on these average reliabilities of raw measures as well.

We plan to supplement/potentially replace this with an additional labeling procedure intending to capture the underlying putative cognitive processes but are deferring this until at least for another week. This exercise might not even prove helpful but at least a per task summary seemed a reasonable first pass.

```{r echo=FALSE}
measure_labels <- read.csv('/Users/zeynepenkavi/Downloads/retest_measure_labels.csv')
measure_labels = measure_labels %>% select(-Patrick_hates)

trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

We filter out the DDM parameters and measures for specific contrasts. Note that this does leave some tasks with measures that are model fits and/or for specific conditions (because at least the current datasets do not include measures that are based on all the trials **and** are raw though I could dive in to variables_exhaustive for such measures. For example the average relialibility for Kirby is based on three discount rates for specific conditions.). Here's the order of tasks by mean reliability sorted for ICC and then Spearman's $\rho$.

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -measure_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(label_1 %in% c('difference', 'EZ', 'hddm') ==FALSE & label_2 %in% c('difference', 'EZ', 'hddm') == FALSE) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  group_by(task_name) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n(),
            num_trials = unique(num_all_trials)) %>%
  arrange(-median_icc, -median_spearman)

tmp %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc'), digits=3)
```

Does number of items in a task have a significant effect on the average ICC of (mostly) raw measures for all trials from a task? No.

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_trials, data = tmp))
```

Does number of items in a task have a significant effect on the average Spearman $\rho$ of (mostly) raw measures for all trials from a task? No.

```{r warning=FALSE, message=FALSE}
summary(lm(median_spearman ~ num_trials, data = tmp))
```

Does number of measures that went in to this reliability estimate have a significant effect on the average ICC of (mostly) raw measures for all trials from a task? It seems the more measures that goes in to summarizing overall reliability for a task the worse it gets. This would make sense if these variables are intended to measure different things and some are less reliable than others.

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_measures, data = tmp))
```

Does number of measures that went in to this reliability estimate have a significant effect on the average ICC of (mostly) raw measures for all trials from a task?  
The same is true for Spearman's correlations as well.

```{r warning=FALSE, message=FALSE}
summary(lm(median_spearman ~ num_measures, data = tmp))
```

In addition to average relilibilities for each task we can look at relilibities of measures organized by the putative cognitive process they are intended to measure. Note however that these labels are not determined in a data driven manner.

```{r warning=FALSE, message=FALSE}
measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  group_by(measure_type) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n()) %>%
  arrange(-median_icc, -median_spearman) %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc'), digits=3)
```

Here is what the distributions look like.  
- Alerting: This labels consists only of ANT alerting contrast (no cue - double cue trials). 4/6 are drift diffusion parameters.    
- Cue encoding: This label consists only of three by two task switch cost contrast (stay (where cue and task remains the same) - switch (where task is the same but cue changes) trials). Again most are drift diffusion parameters.  
- Delay discounting: This label consists of the discount rates for the Bickel titrator (3), Kirby items (3) and discount titrate (model fit to random set of stimuli). **Does the discrepancy between ICC and Spearman for these measures, given the specific interest in these tasks in the literature warrant a more detailed look in to these?**  
- Information use: This label includes Columbia Card Task (both versions) and information sampling task.  
- Inhibition: This label includes go no-go, motor selective stop signal, stimulus selective stop signal and regular stop signal (excluding specific constrasts).  
- Intelligence: This label includes ravens and tower of london.  
- Learning: This label includes ART, CCT, two stage decision, dietary decision task, hierarchical rule task, information sampling task, probabilistic selection task, shift task, two stage task.   
- Neutral writing: This label consists only of the writing task.  
- Orienting: This labels consists only of ANT orienting contrast (center - spatial cue trials). 4/6 are drift diffusion parameters.   
- Positive writing: This label consists only of the writing task.  
- Proactive control: This labels consists of one DPX contrast (AY - BY trials), motor selective stop signal and regular stop signal.   
- Reactive control: This labels consists of one DPX contrast (BX - BY trials) and motor selective stop signal.  
- Resisting proactive interference: This label consists of one contrast in directed forgetting and recent probes tasks (negative (where the stimulus to respond to is in forgetting set) and control (where the stimulus to respond to is new) trials).  
- Response conflict: This label consists of ANT conflict contrast (congruent-incongruent), DPX, local global conflict contrast, shape matching, simon, stroop.  
- Response selection: This label consists of choice reaction time task.  
- Risk Taking: This label consists of ART and CCT.  
- Selective attention: This label consists of local global task and ANT (all trials).  
- Set shifting: This label consists of local global, PRT and three by two.    
- Speed: This label consists of raw response times in information sampling task, simple reaction time, tower of london.  
- Working memory: This label consists of adaptive n back, digit span, directed forgetting, keep track, recent probes, spatial span.

```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  select(measure_type, icc, spearman) %>%
  gather(key, value, -measure_type) %>%
  ggplot(aes(value, fill=key))+
  geom_histogram(position='identity', alpha=0.5)+
  theme_bw()+
  facet_wrap(~measure_type)+
  theme(legend.title = element_blank())
```

### Surveys

>>>>

---
title: 'SRO Retest Data Cleaning'
output:
github_document:
toc: yes
toc_float: yes
---

#################################
NOT USED
#################################

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(psych)
library(rmarkdown)
library(psych)
library(stringr)
library(plotly)
library(DT)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
render_this <- function(){rmarkdown::render('SRO_Report.Rmd', output_dir = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/reports', html_notebook(toc = T, toc_float = T, code_folding = 'hide'))}
options(scipen = 1, digits = 4)
```

@@
